The Bot with the Eloquent Speech: Exploring the Relationship between Perceptions of Conversational Agents and Their Conversation Architecture

### 2023-02-14 paper.tex
second pass up to 3.1

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Previous research has demonstrated a relationship between the structure of conversational agents' natural language interaction and how these agents are perceived by their users. However, the systematic connections between the perception of agents and elements of conversation architecture are not well understood. Furthermore, there is no clear definition of the aspects of perception that are affected and the elements of conversation architecture. In this paper, we present a synthesized framework demonstrating the connections between perceptions of agents and conversation architecture elements, using a taxonomy of eleven perceptions grouped into four categories, and ten elements grouped into three categories. We highlight the need for further research for the perception of agent's ability and perception of social connection with agent's, especially as it relates to the linguistic style of conversational agents.  Also, there is the need to be consistent in the assessment of agents' perceptions, and address ethical concerns in design.
\end{abstract}

@@

Conversational agents (CAs) are software-based systems, either voice of text-based, designed to interact with humans using natural language. Adoption for CAs such as Alexa (Amazon) or Siri (Apple) are rapidly growing in the market, with half of US internet users own one or more smart speaker devices \cite{2022comscore}. These conversational agents are designed to be ubiquitous, easily accessible through smartphones, tablets, laptops and smart devices. Because these agents leverage natural language for interaction, they solicit social responses from users, encouraging users to attribute lifelike qualities, a.k.a. anthropomorphized perceptions to CAs \cite{eyssel2012if}.

Anthropomorphism is defined as the attribution of traits that we typically associate with being distinctly human to nonhuman entities \cite{waytz2010sees}. For example, users tend to personify conversational agents like Amazon Alexa by using person pronouns instead of object pronouns in online reviews \cite{purington2017alexa}.  Research shows that anthropomorphic design is beneficial to establishing and maintaining trust between users and conversational agents \cite{seeger2021chatbots}. Also, speakers tend to align their lexical choices with conversational agents, similar to human-human conversations \cite{cowan2015does}. Anthropomorphism has been shown to ease user interactions, e.g. making agents more approachable, engaging, and trustworthy when they exhibit human characteristics \cite{qiu2009evaluating}. These processes are of particular interest for human-agent communication because they form the basis for user expectations and predispositions regarding agents' behavior, reliability of the information, etc. \cite{kuzminykh2020genie}.

One of the key factors for anthropomorphized perception is how conversational agents convey information through verbal or non-verbal cues. Studies have found that conversational agents are perceived as more socially present and emotionally intelligent if they use affect language such as sentiment-adaptive response based on user's utterances, resulting in higher user satisfaction \cite{diederich2019emulating}\cite{yang2017perceived}. Also, CAs that mimic user behaviour through lexical alignment reduce users' cognitive workload, resulting in higher engagement with the interaction \cite{spillner2021talk}. Non-linguistic cues such as expressive prosody contribute to higher perceived intimacy with the user, as well as higher perceived enjoyment and ease of use \cite{kim2020can}. While there is existing research on the impact of conversation architecture on anthropomorphized perceptions of agents, their findings are hard to generalize due to obstacles in consistency of taxonomy and measures.
%However, conversational architecture design may lead to negative consequences, such as an agent with an extroverted personality may be seen as less trustworthy, as users are irritated by the system's overly familiar attitude \cite{andrews2012system}.

 There have been recent efforts to systematically analyze UX measurements used to evaluate conversational agents. In Clark et al's review for the state of speech interfaces, the authors created a list the different concepts used to measure disembodied voice-based agents such as user attitudes and task performance \cite{clark2019state}. Rapp et al. \cite{rapp2021human} analyzed literature related specifically to text-based CAs, where the themes emerged from their analysis has aspects related to the anthropomorphized perceptions, like humanness and emotional experience. Most recently, Zheng et al's literature review on the UX research of CAs \cite{zheng2022ux} categorized the metrics based on the type of method used, including technical measures like system error rate and perception measures like perceived social presence. Specifically to measures related to anthropomorphized perceptions, there is a gap of research to systematically analyze the different aspects of these perceptions, as well as a consistent taxonomy to refer to them. Similarly for conversation architecture elements, there is existing research to systematically analyze them, like Feine et al \cite{feine2019taxonomy}'s overall framework provides consistent categories for social cues. However, it is unknown which conversation architecture elements have impacts on the perceptions of agents. Lastly, while there are various studies investigating the effect of  conversational architecture elements on the anthropomorphized perception of agents, we did not find any systematic review and analysis to generalize the findings into a usable framework.

We are addressing this research gap by analyzing how to design conversation architecture in order to orchestrate anthropomorphized perceptions of conversational agents. First, we want to understand (RQ1) what aspects of anthropomorphized perceptions are being explored that are affected by conversation architecture. Conversation architecture has various aspects that affect a CA's utterances such as the use of humour, emoticons, or lexical alignment. This leads to our second research question: (RQ2) what are the elements of conversation architecture that are relevant to anthropomorphized perceptions? Lastly, we want to understand the impact of conversation architecture on perceptions of agent, which leads to our third research question: (RQ3) what are the relationships between the specifics of conversation architecture elements and the anthropomorphized perceptions of CAs?

To answer these questions, we need to synthesize existing literature by performing a comprehensive review from 2010 to 2022 on the impact of conversation architecture elements on anthropomorphized perceptions of CAs (n=57). We collected perceptions that are affected by conversation architecture and grouped them into four categories and eleven themes (Table \ref{tab:perceptions}). Similarly, we analyzed the conversation architecture elements explored in literature and grouped them into three categories and ten themes (Table \ref{tab:cues}). Lastly, we coded the explored connections between conversation architecture and perceptions and demonstrated the impacted relationship into a generalized framework (Figure \ref{fig:heatmap-impact}). Through our analysis, we also found that there are inconsistencies on how anthropomorphized perceptions are measured across studies, as well as influencing factors other than conversation architecture that may be impacting the perceptions of agents.

This work contributes to the HCI community by presenting an in-depth analysis on the impact of conversation architecture on anthropomorphized perceptions. We created a taxonomy for perceptions and conversation architecture elements to provide consistent terms that can be used in future research. Also, our analysis extends existing reviews on UX research of CAs \cite{clark2019state}\cite{rapp2021human}\cite{zheng2022ux} by providing a generalized framework outlining the relationships between conversation architecture and anthropomorphized perceptions. This framework provides the foundational knowledge to design conversation agents to solicit specific social responses from users.

In the remainder of the paper, we first outline our literature review and data analysis process, as well as the characteristics of the papers in our reviewed corpus. Following that, we describe in detail the taxonomy of anthropomorphized perceptions, taxonomy of conversation architecture element, as well as the framework on the relationships between perceptions and conversation architecture. Finally, we discuss the research challenges and opportunities as well as ethical considerations for designing conversational architecture to orchestrate anthropomorphized perceptions of conversational agents.

@@

This section describes the research methodologies for the search procedure, selection criteria and data analysis process.

@@

Following PRISMA guidelines \cite{prisma}, we reviewed and selected literature relevant to the relationships between the specifics of conversation architecture elements and anthropomorphized perceptions of CAs. Searches were carried out in the ACM Digital Library between January 1 and January 15, 2023. During initial analysis, it didn't seem like there were common terms used to refer to either conversation architecture elements, or anthropomorphized perceptions. As such, search terms were kept general to literature related to conversational agents. Based on search terms used in previously published literature reviews \cite{clark2019state}\cite{rapp2021human} , the following keywords have been identified to search for publications related to conversational user interfaces:
\newline

@@

The initial query retrieved 2901 unique publications. The following selection criteria are applied to identify literature related to the effect of conversational architecture on perception of agents:
\begin{itemize}
  \item Peer-reviewed publications written in English
  \item Voice-based or text-based conversational agents
  \item Experiment study on the impact of conversation architecture elements on the anthropomorphized perceptions of agents
  \item Impact of conversation architecture elements can be separated from other effects (e.g. embodiment)

@@

We screened the titles and abstracts of the papers based on the selection criteria below which resulted in 221 papers. Reviewing the full articles of these publication resulted in 49 papers for analysis. An additional search for related literature was also performed to find new publications, adding 8 more papers to the corpus. We identified 57 relevant articles that met our selection criteria (see Figure \ref{fig:prisma}).

@@

On the modality characteristics of the conversational agents in the corpus, there is a roughly even split between the modalities, with slightly more papers (n=30) studying voice-based CAs, and the rest (n=27) studying text-based CAs.

@@

\subsubsection*{Anthropomorphized Perceptions} To study RQ1, the exact quotes of anthropomorphized perception measurements were extracted from each paper. Inductive thematic analysis was used to group these verbatim quotes into codes based on their meanings. For example, the code likeability of the agent contains the questionnaire item "this voice agent was likeable" used by \cite{cuadra2021my}\cmt{[67]}, Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed}'s set of questions on likeability used by \cite{linnemann2018can}\cmt{[15]}, and Subjective Assessment of Speech System Interface (SASSI) questionnaire \cite{hone2000towards}\cmt{sassi}'s set of questions on likeability used by  \cite{chan2021kinvoices}\cmt{[74]}\cite{choi2020nobody}\cmt{[54]}. This step resulted in 83 unique codes. The codes are then organized based similarity of the meanings, resulting in 11 basic themes. For example, the basic theme of \textit{personality traits} of the agent contains measurements such as likeability, friendliness, and warmth. Lastly, the 11 basic themes are categorized into 4 organizing themes: \textit{perception of interaction with agent, perception of agent's ability, perception of social connection with agent, and perception of agent's humanness} (Table \ref{tab:perceptions}).

\subsubsection*{Conversation Architecture Elements}
The study RQ2, the conversation architecture elements used in the study were extracted from each paper. Some of the papers were applying multiple elements to the CA design, such as the the anthropomorphic design used in this study \cite{seeger2021chatbots}\cmt{[35]}. In the case where multiple elements were used, the codes broke down the conversation architecture into its core components. For example, the anthropomorphic design used in Seeger et al's study \cite{seeger2021chatbots}\cmt{[35]} was broken down to the elements of emotional expressions, is-typing indicator, emoticons, and response delay to capture as codes. 58 unique codes were created as part of this process, capturing elements like sentiment-adaptive responses \cite{diederich2019emulating}\cmt{[25]}, lexical alignment \cite{spillner2021talk}\cmt{[18]}, and typos \cite{westerman2019believe}\cmt{[9]}. Codes were then organized based on similarity of the elements. For example, the basic theme of \textit{disfluency} contains elements of fillers \cite{jeong2019exploring}\cmt{[10]}\cite{wester2015artificial}\cmt{[14]}, interjections \cite{ceha2022expressive}\cmt{[77]}\cite{hu2021enhancing}\cmt{[56]}, and repetitions \cite{yang2021effect}\cmt{[72]}. Lastly, these basic themes are then categorized into organizing themes adapted from the social cues taxonomy structure by Feine et al \cite{feine2019taxonomy}. Based on this paper, the Verbal-Content category is mapped to our organizing theme of \textit{linguistic content}, and the Verbal-Style category is mapped to our organizing theme of \textit{linguistic style}.
Given our analysis focuses on the text and voice aspects of a conversational agent and does not analyze the embodiment aspects of design, the visual category is not applicable for our analysis. Also, we collapsed the auditory and invisible categories into the organizing theme of \textit{non-linguistic format} (Table \ref{tab:cues}).

\subsubsection*{Relationship Between Anthropomorphized Perceptions and Conversation Architecture Elements}
For RQ3, we extracted the relationships between perceptions and architecture elements from each paper based on the codes developed for anthropomorphized perceptions and conversation architecture elements above, noting whether the architecture element had an impact of on the perceptions. Out of the 57 papers, 264 connections between anthropomorphized perceptions and conversation architecture elements were found. To analyze specifically the relationships between the specifics of conversation architecture that affect the anthropomorphized perceptions of CAs, 69 connections that did not result in significant relationships were discarded, resulting in 195 relationships for analysis. For example, the specific connection between style matching and user satisfaction was discarded because Hoegen et al. \cite{hoegen2019end}\cmt{[31]} did not find a significant difference between the style matching agent and the non-style matching agent for overall interaction satisfaction. Out of the 195 relationships, 185 were relationships between individual architecture elements and perceptions. These single architecture element to perception relationships are visualized as a heatmap based on literature coverage within the literature review corpus, as seen in Figure \ref{fig:heatmap-impact}. Perceptions impacted by the composite of multiple architecture elements in an agent will be discussed in more details in the Composite Architecture Elements section.

@@

As shown in Table \ref{tab:perceptions}, there are four organizing themes defined for the anthropomorphized perceptions of conversational agents covering 11 basic themes of perception. The details of each theme are discussed below.

\textbf{Perception of interaction with agent} evaluates the overall interaction quality users had with the conversational agent. The three basic themes for this perceptions are usability, engagement and satisfaction. \textit{Usability} measures the utilitarian aspects of the interaction, whether it was accurate, easy to use, efficient or helpful. Some commonly used methods to measure usability include the response accuracy portion of the SASSI questionnaire \cite{hone2000towards}\cmt{sassi}, or the NASA Task Load Index (NASA-TLX) \cite{hart1988development}\cmt{nasa} to measure cognitive workload. Questions such as "the system is easy to use" or "it is easy to understand the agent" are also measures of usability. \textit{Engagement} on the other hand measures users' emotional reactions, whether they have enjoyed the conversation with CA, or annoyed or frustrated with the interaction. Some commonly used methods to measure engagement include  the annoyance portion of the SASSI questionnaire \cite{hone2000towards}\cmt{sassi}, and the Use Engagement Scale (UES) \cite{o2018practical}\cmt{ues}. Questions such as "I enjoyed using the system" or "I felt frustrated with the agent" are also measures of engagement. Lastly, \textit{Satisfaction} measures users' overall satisfaction interacting with the agent. Questions such as "the overall assessment of conversing with the CA was satisfactory" are used to measure this perception.

\textbf{Perception of agent's ability} evaluates the perceived capabilities of the agent. While there are some elements of system performance in this measure, the main aspect we are interested in is the perception differences for agents with identify system capabilities but with different architecture elements. Specifically, \textit{intelligence} measures the agent's expertise and knowledge, and it is commonly administered through the Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed} on the set of questions related to perceived intelligence. Also survey questions such as asking users about the agent's intelligence and domain knowledge are used to measure perceived intelligence. \textit{Competence} takes intelligence one step further by evaluating the agent's ability to put its intelligence into practice. To measure a CA's competence, there are usually surveys or qualitative feedback that the agent is capable and competent, and that the users have confidence in the agent's ability to get the job done. Lastly, \textit{trust} evaluates agent's truthfulness, credibility and benevolence. Some commonly used methods to evaluate trust include the Trust Propensity Scale \cite{mayer1999effect} and the Individualized Trust Scale (ITS) \cite{wheeless1977measurement}. Questions such as "is the agent honest" and "can I trust the agent with sensitive information" are also measures of trust.

\textbf{Perception of the social connection with agent} evaluates the emotional connections that users have with agents. The three basic themes of conversation tone, social presence and intimacy are included in this category. \textit{Conversation tone} measure the affective impression of the agent's tone, such as empathy, expressiveness, or passionate. Also, it measures whether the tone used by the agent is persuasive or appropriate for the conversation. \textit{Social presence} measures the sense of connectness and psychological distance users has with an agent. Some aspects in this measure include a sense of familiarity or similarity with the agent, whether users feel the agent behaves like them or have similar attitudes to them. The measure of \textit{Intimacy} extends social presence into the realm of the quality of relationships with an agent. Some commonly used measures include the set of social attraction questions from the interpersonal attraction questionnaire \cite{mccroskey1975development} and the quality of relationship inventory (QRI) \cite{pierce1997assessing}. These questionnaires include questions like "I think the agent could be a friend of mine", or "I feel we could establish a personal relationship with each other".

\textbf{Perception of agent's humanness} evaluates human-likeness and the anthropomorphized personality traits that users assign to agents. \textit{Human-likeness} measures whether the agent presented itself as natural and human-like, or artificial and machine-like. The Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed} set of questions related to anthropomorphism and the Ascent of Man scale \cite{kteily2015ascent} are commonly used methods to assess human-likeness. Survey questions with semantic scales such as "human-like / machine-like" and "artificial / natural" are also used for this perception. \textit{Personality traits} captures the human characteristics that are attributes to the agent, such as warm, friendly, likeable, and polite. This is commonly captured in the qualitative feedback from users, commenting on whether the agent is extroverted or introverted, or the human characteristics they perceive the agent such as funny or witty. Some studies used the measure of the Big-5 personality traits \cite{gosling2003very} to map an agent's disposition on various personality dimensions.

Overall there is good coverage of perception measures based on the list of papers reviewed in our corpus. The most commonly measured perceptions are related to the interaction with agent, followed by the perception of agent's humanness. There are two basic themes that are the least measured compared to the others: competence under the perception of agent's ability, and intimacy under the perception of social connection with agent. This may be due to the controlled lab settings for the experiments, where participants are given the scenarios for interaction. This environment is not conducive to forming relationships with a conversational partner, as noted by Linnerman et al. in their discussions \cite{linnemann2018can}\cmt{[15]}. Also, the same factor could impact the assessment of agent's abilities, as the users may not feel like they have the expertise to assess an agent's competence. These factors could contribute to the reasons why competency and intimacy are not as frequently measured in experimental studies.

% Add citations

@@

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Previous research has demonstrated a relationship between the structure of conversational agents' natural language interaction and how these agents are perceived by their users. However, the systematic connections between the perception of agents and elements of conversation architecture are not well understood. Furthermore, there is no clear definition of the aspects of perception that are affected and the elements of conversation architecture. In this paper, we present a synthesized framework demonstrating the connections between perceptions of agents and conversation architecture elements, using a taxonomy of eleven perceptions grouped into four categories, and ten elements grouped into three categories. We highlight the need for further research for the perception of agent's ability and perception of social connection with agent's, especially as it relates to the linguistic style of conversational agents.  Also, there is the need to be consistent in the assessment of agents' perceptions, and consider ethical concerns in design.
\end{abstract}

@@

Conversational agents (CAs) are designed to interact with humans using natural language through voice or text modalities. Adoption for CAs such as Alexa (Amazon) or Siri (Apple) are rapidly growing in the market, with half of Internet users in United States own one or more smart speaker devices \cite{2022comscore}. These conversational agents are designed to be ubiquitous, easily accessible through smartphones, tablets, laptops and smart devices. Since these agents interact with users using natural language, there are nuances in their communications affected by conversation architecture, which is what they say and how they say it.
%they solicit social responses from users, encouraging users to attribute lifelike qualities, a.k.a. anthropomorphized perceptions to CAs \cite{eyssel2012if}.

%Anthropomorphism is defined as the attribution of traits that we typically associate with being distinctly human to nonhuman entities \cite{waytz2010sees}. For example, users tend to personify conversational agents like Amazon Alexa by using person pronouns instead of object pronouns in online reviews \cite{purington2017alexa}.  Research shows that anthropomorphic design is beneficial to establishing and maintaining trust between users and conversational agents \cite{seeger2021chatbots}. Also, speakers tend to align their lexical choices with conversational agents, similar to human-human conversations \cite{cowan2015does}. Anthropomorphism has been shown to ease user interactions, e.g. making agents more approachable, engaging, and trustworthy when they exhibit human characteristics \cite{qiu2009evaluating}. These processes are of particular interest for human-agent communication because they form the basis for user expectations and predispositions regarding agents' behavior, reliability of the information, etc. \cite{kuzminykh2020genie}.

Previous research has show that there is a connection between conversation architecture and the perception of agents. Studies have found that conversational agents are perceived as more socially present and emotionally intelligent if they use affect language such as sentiment-adaptive response based on user's utterances, resulting in higher user satisfaction \cite{diederich2019emulating}\cite{yang2017perceived}. Also, CAs that mimic user's utterances through lexical alignment reduce users' cognitive workload, resulting in higher engagement with the interaction \cite{spillner2021talk}. Non-linguistic cues such as expressive prosody contribute to higher perceived intimacy with the user, as well as higher perceived enjoyment and ease of use \cite{kim2020can}. Although research on how conversation architecture affects agents' perceptions already exists, there is a lack of understanding on the coverage of relationships explored in literature, as well as synthesizing these findings into a generalized framework. One of the possible issues is the lack of clear and consistent definition on the aspects of user's perceptions of agent and conversation architectural elements.
%However, conversational architecture design may lead to negative consequences, such as an agent with an extroverted personality may be seen as less trustworthy, as users are irritated by the system's overly familiar attitude \cite{andrews2012system}.

Related to the perception aspects of agents, there have been recent efforts to systematically analyze user experience criteria used to evaluate conversational agents. Some review papers have created a list of dimensions used to assess conversational agents, which include the perceptions of agents (e.g. empathy, humanness) alongside other assessments (e.g. physiological data) \cite{clark2019state}\cite{finch2020towards}.
%In Clark et al's review for the state of speech interfaces, the authors created a list the different concepts used to assess disembodied voice-based agents such as user attitudes and task performance \cite{clark2019state}. 
Other papers categorized CA assessments into themes. For example, Rapp et al. \cite{rapp2021human} generated themes related to how users' experience with chatbots such as experiencing the chatbot including assessments on expectation, perception, satisfaction etc. Most recently, Zheng et al's literature review on the UX research of CAs \cite{zheng2022ux} categorized the metrics based on the type of method used, including technical measures like system error rate and perception measures like perceived social presence. However, existing studies use different terminologies and categories across aspects of perception. As an illustration, it is unclear if the theme of experiencing the chatbot from Rapp et al. \cite{rapp2021human} contains similar aspects to the perception towards CA in Zheng et al \cite{zheng2022ux}'s analysis.

For conversation architecture elements, Feine et al \cite{feine2019taxonomy} created a foundational knowledge by providing a taxonomy of social cues, which is defined as "a cue of a CA that triggers a social reaction of the user towards the CA." Elements of conversation architecture are captured in the taxonomy, alongside other social cues such as the embodied agents' appearance and movements. Clark et al \cite{clark2019state} had also captured conversation architecture elements under the research topic of system speech production, with similar categorizations across content, style, and other topics. While these papers captured conversation architecture elements as part of their broader scope of research, there exist the need for standardized definitions specifically for conversation architecture that reflects its unique characteristics.

%there is existing research to systematically analyze them, like Feine et al \cite{feine2019taxonomy}'s overall framework provides consistent categories for social cues. However, it is unknown which conversation architecture elements have impacts on the perceptions of agents. Lastly, while there are various studies investigating the effect of  conversational architecture elements on the anthropomorphized perception of agents, we did not find any systematic review and analysis to generalize the findings into a usable framework.

In this paper, we are addressing these research gaps by analyzing how to design conversation architecture in order to orchestrate the perceptions of conversational agents. First, we want to understand \textbf{(RQ1) what perception aspects of agents are being explored in relation to the effects of conversation architecture elements}. Various elements of conversation architecture can be used in a CA's utterances such as humour, emoticons, or lexical alignment. This leads to our second research question: \textbf{(RQ2) what elements of conversation architecture are relevant to the perceptions of agents?} Lastly, we want to understand the impact of conversation architecture on perceptions of agent, which leads to our third research question: \textbf{(RQ3) what are the relationships between the specifics of conversation architecture elements and the perceptions of CAs?}

To answer these questions, we performed a comprehensive review of existing literature from 2010 to 2022 and synthesized findings on the impact of conversation architecture elements on the perceptions of CAs based on relevant papers (n=57). We collected perceptions of agents that are affected by conversation architecture,  grouped the aspects into four categories and eleven sub-categories (Table \ref{tab:perceptions}). We also analyzed conversation architecture related to perception of agents and grouped them into three categories and ten elements (Table \ref{tab:cues}). Lastly, we coded the explored connections between conversation architecture and perceptions and demonstrated the relationship into a synthesized framework (Figure \ref{fig:heatmap-impact}). Through our analysis, we also found that there are inconsistencies in the assessment of perceptions across studies, as well as influencing factors other than conversation architecture that impact the perceptions of agents.

This work contributes to the HCI community by presenting an in-depth analysis on the affect of conversation architecture on perceptions of conversational agents. We created a taxonomy for perception aspects and conversation architecture elements, providing clear and consistent terms that can be used in future research. Also, our analysis extends existing reviews on UX research of CAs \cite{clark2019state}\cite{rapp2021human}\cite{zheng2022ux} by providing a synthesized framework outlining the relationships between conversation architecture and perceptions of agents. With the help of this framework, we are taking the first step to understand how to design conversational agents to solicit specific social reactions from users.

In the remainder of the paper, we first outline our data collection and data analysis process, as well as the characteristics of the papers in our reviewed corpus. Following that, we describe in detail the taxonomy for the perception of agents, the taxonomy for conversation architecture element, as well as the framework on the relationships between perceptions of agents and conversation architecture. Finally, we discuss research challenges and opportunities as well as ethical considerations for designing conversational architecture to orchestrate perceptions of conversational agents.

@@

This section describes the research methodologies used for data collection and data analysis. Additionally, it presents the characteristics of the papers in our reviewed corpus.

@@

Following PRISMA guidelines \cite{prisma}, we reviewed and selected literature exploring the relationship between the specifics of conversation architecture elements and perceptions of CAs. Searches were carried out in the ACM Digital Library between January 1 and January 15, 2023. Through our analysis, we didn't find any common terms used to refer to either conversation architecture elements, or aspects of perceptions. As such, we kept our search term to capture articles related to conversational agents. Based on search terms used in previously published literature reviews \cite{clark2019state}\cite{rapp2021human}, the following keywords are used to search for publications related to conversational user interfaces:
\newline

@@

The following selection criteria are applied to identify literature related to the effect of conversational architecture on the perception of agents:
\begin{itemize}
  \item The paper was published between 2010 and 2022.
  \item The paper is peer-reviewed and written in English.
  \item The paper contains the use of voice-based or text-based conversational agents.
  \item The paper contains studies on the impact of conversation architecture elements on the perceptions of agents.
  \item The paper contains impact of conversation architecture elements can be separated from other effects (e.g. embodiment).

@@

The initial query retrieved 2901 unique publications. We screened the titles and abstracts of the papers based on the selection criteria above, resulting in 221 papers. The fully body of these papers are then reviewed, selecting papers that met the criteria. A search for additional literature was also performed to supplement the corpus, adding 8 more papers to the selection. In total, we identified 57 relevant articles for analysis (see Figure \ref{fig:prisma}).

@@

On the modality characteristics of the conversational agents in the corpus, there is a roughly even split between the modalities, with slightly more papers exploring voice-based CAs (n=30) compared to text-based CAs (n=27).

@@

\subsubsection*{Perceptions of Agents} To understand the perceptions of agents related to conversation architecture (RQ1), assessments for the perceptions of agents were extracted from each paper. We used inductive thematic analysis to group these assessments into codes based on their meanings. For example, the code likeability of the agent contains assessments such as the questionnaire item "this voice agent was likeable" used by \cite{cuadra2021my}\cmt{[67]}, Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed}'s set of questions on likeability used by \cite{linnemann2018can}\cmt{[15]}, and Subjective Assessment of Speech System Interface (SASSI) questionnaire \cite{hone2000towards}\cmt{sassi}'s set of questions on likeability used by \cite{chan2021kinvoices}\cmt{[74]}\cite{choi2020nobody}\cmt{[54]}. This step resulted in 83 unique codes. These codes are then organized based similarity of the meanings, resulting in 11 aspects of perceptions. For example, the aspect of \textit{personality traits} of the agent contains measurements such as likeability, friendliness, and warmth. Lastly, the 11 aspects are grouped into 4 categories: \textit{perception of interaction with agent, perception of agent's ability, perception of social connection with agent, and perception of agent's humanness} (Table \ref{tab:perceptions}).

\subsubsection*{Conversation Architecture Elements}
As to the elements of conversation architecture that are relevant to the perception of agents (RQ2), we extracted the conversation architecture used in each paper as codes. Some of the papers used multiple features in their CA design, such as the the anthropomorphic agent used in Seeger et al's study \cite{seeger2021chatbots}\cmt{[35]}. In these composite situations, we broke them down into their individual features. For example, the anthropomorphic agent \cite{seeger2021chatbots}\cmt{[35]} was separated into emotional expressions, is-typing indicator, emoticons, and response delay and captured as codes. 58 unique codes were created as part of this process, capturing features like sentiment-adaptive responses \cite{diederich2019emulating}\cmt{[25]}, lexical alignment \cite{spillner2021talk}\cmt{[18]}, and typos \cite{westerman2019believe}\cmt{[9]}. We then organized these codes into elements based on similarity. For example, the element of \textit{disfluency} contains fillers \cite{jeong2019exploring}\cmt{[10]}\cite{wester2015artificial}\cmt{[14]}, interjections \cite{ceha2022expressive}\cmt{[77]}\cite{hu2021enhancing}\cmt{[56]}, and repetitions \cite{yang2021effect}\cmt{[72]}. Lastly, these elements are grouped into categories adapted from the social cues taxonomy structure by Feine et al \cite{feine2019taxonomy}. Based on this paper, the Verbal-Content category is mapped to our category of \textit{linguistic content}, and the Verbal-Style category is mapped to our category of \textit{linguistic style}. Given our analysis focuses on the conversational aspects of a CA, the social cues' categories related to embodiment aspects were not used. Also, we collapsed the auditory and invisible categories into our category of \textit{non-linguistic format} (Table \ref{tab:cues}).

\subsubsection*{Relationship Between Perceptions of Agents and Conversation Architecture}
To study the relationship between conversation architecture and conversation architecture, we extracted the connections explored in each paper, using the perception aspects and conversation architecture elements developed in the previous data analysis steps. Each connection's effect was also recorded based on whether an impact was found in the study and the nature of that impact. Out of our review corpus, 265 explored connections between perceptions of agents and conversation architecture were found. To analyze specifically the relationships between the specifics of conversation architecture that affect the perceptions of CAs, 72 connections that did not result in significant relationships were discarded, resulting in 193 relationships for analysis. For example, the connection between the matching style of a CA and user satisfaction was discarded because Hoegen et al. \cite{hoegen2019end}\cmt{[31]} did not find a significant difference between the style matching agent and the non-style matching agent for overall interaction satisfaction. Out of the 193 relationships, 10 were relationships using agents using multiple architecture elements. They were not included in the framework as it is not possible attribute the perceptions of agents to its underlying elements. Overall, 183 relationships between individual architecture elements and perceptions of agents are incorporated into our synthesized framework, visualized as a heatmap to demonstrate their relationships with each other (Figure \ref{fig:heatmap-impact}).

%183 were relationships between individual architecture elements and perceptions. These single architecture element are included in our synthesized frame, 
%CAs using multiple conversation architecture elements were discarded because perceptions are measured at the aggregate level, we cannot attribute the perceptions of agents down to its underlying elements.
%Perceptions impacted by the composite of multiple architecture elements in an agent will be discussed in more details in the Composite Architecture Elements section.

@@

As shown in Table \ref{tab:perceptions}, there are four categories defined for the perceptions of conversational agents covering 11 aspects of perception. The details of each aspect are discussed below.

\textbf{Perception of interaction with agent} assesses the overall interaction quality users had with the conversational agent. The three aspects for this perceptions are usability, engagement and satisfaction. \textit{Usability} captures the utilitarian aspects of the interaction, whether it was accurate, easy to use, efficient or helpful. Some commonly used methods to measure usability include the response accuracy portion of the SASSI questionnaire \cite{hone2000towards}\cmt{sassi}, or the NASA Task Load Index (NASA-TLX) \cite{hart1988development}\cmt{nasa} to measure cognitive workload. Questions such as "the system is easy to use" or "it is easy to understand the agent" are also used to evaluate usability. \textit{Engagement} on the other hand captures users' emotional reactions to the CA, whether they have enjoyed the conversation, or annoyed or frustrated with the interaction. Some commonly used methods to measure engagement include the annoyance portion of the SASSI questionnaire \cite{hone2000towards}\cmt{sassi}, and the Use Engagement Scale (UES) \cite{o2018practical}\cmt{ues}. Questions such as "I enjoyed using the system" or "I felt frustrated with the agent" are also used to evaluate engagement. Lastly, \textit{Satisfaction} captures users' overall satisfaction interacting with the agent. Questions such as "the overall assessment of conversing with the CA was satisfactory" are used to evaluate this perception.

\textbf{Perception of agent's ability} assesses the perceived capabilities of the agent. While there are some elements of system performance in this category, we are mainly concerned with how agents with varied architecture elements but equivalent system capabilities are perceived differently. Specifically, \textit{intelligence} captures the agent's expertise and knowledge, and it is commonly administered through the Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed} using the set of questions related to perceived intelligence. Also survey questions such as asking users about the agent's intelligence and domain knowledge are used to evaluate perceived intelligence. \textit{Competence} takes intelligence one step further by examining the agent's ability to put its intelligence into practice. Surveys or qualitative feedback are usually used to assess if the agent is capable and competent, and whether users have confidence in the agent's ability to get the job done. Lastly, \textit{trust} captures agent's truthfulness, credibility and benevolence. Some commonly used methods to evaluate trust include the Trust Propensity Scale \cite{mayer1999effect} and the Individualized Trust Scale (ITS) \cite{wheeless1977measurement}. Questions such as "is the agent honest" and "can I trust the agent with sensitive information" are also used to evaluate trust.

\textbf{Perception of the social connection with agent} assesses the emotional connections that users have with agents. This category includes perception aspects of conversation tone, social presence and intimacy. \textit{Conversation tone} captures the affective impression of the agent's tone, such as empathy and expressiveness. Also, this aspect includes whether the tone used by the agent is persuasive or appropriate for the conversation. \textit{Social presence} captures the sense of connectness and psychological distance users have with an agent. Some aspects in this category include the sense of familiarity or similarity with the agent, and whether users feel the agent behaves like them or have similar attitudes to them. The category of \textit{Intimacy} extends social presence into the realm of the quality of relationships with an agent. Some commonly used methods to assess intimacy include the set of social attraction questions from the interpersonal attraction questionnaire \cite{mccroskey1975development} and the quality of relationship inventory (QRI) \cite{pierce1997assessing}. These questionnaires include questions like "I think the agent could be a friend of mine", or "I feel we could establish a personal relationship with each other".

\textbf{Perception of agent's humanness} assesses the specifics of anthropomorphized perceptions of conversational agents. \textit{Human-likeness} captures whether the agent presented itself as natural and human-like, or artificial and machine-like. The Godspeed questionnaire \cite{bartneck2009measurement}\cmt{godspeed} set of questions related to anthropomorphism and the Ascent of Man scale \cite{kteily2015ascent} are commonly used methods to evaluate human-likeness. Survey questions with semantic scales such as "human-like / machine-like" and "artificial / natural" are also commonly used for this aspect of perception. \textit{Personality traits} captures the human characteristics that are attributes to the agent, such as warm, friendly, likeable, and polite. This is commonly collected as qualitative feedback from users, commenting on whether the agent is extroverted or introverted, or the human characteristics they perceive the agent such as funny or witty. Sometimes the Big-5 personality traits questionnaire \cite{gosling2003very} is used to map an agent's disposition on various personality dimensions.

Overall there is a good coverage of different aspects of perception based on the list of papers reviewed in our corpus. The most commonly explored aspect categories is the perception the interaction with agent, followed by the perception of agent's humanness. There are two aspects that are the least explored compared to the others relationships: competence within the perception of agent's ability, and intimacy within the perception of social connection with agent. This may be due to the controlled lab settings for the experiments, where participants are given the scenarios for interaction. This type of environment is not conducive to forming relationships with a conversational partner, as noted by Linnerman et al. in their discussions \cite{linnemann2018can}\cmt{[15]}. The same factor could impact the assessment of agent's competence, as the users may not feel like they have the expertise to assess their perceptions. 
%These factors could contribute to the reasons why competency and intimacy are not as frequently measured in experimental studies.

@@

%CW - second pass up to here
